Report for Tissue-Specific Proteins

### 2021-10-07 SeldinReport.Rmd
havent committed in a while

@@

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_liver = read.csv("Liver_WGCNA.csv")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")
sec_annot = read.csv("Sec_Annotation.csv")
tissuespec_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
tissuespec_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
ts_protein_brain = read.csv("TissueSpecific2_Protein_Brain.csv")
ts_protein_liver = read.csv("TissueSpecific2_Protein_Liver.csv")
tsns_protein_brain = read.csv("NSTissueSpecific_Protein_Brain.csv")
tsns_protein_liver = read.csv("NSTissueSpecific_Protein_Liver.csv")

@@

protein_liver = fix.read.csv(protein_liver)
protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)
sec_annot = fix.read.csv(sec_annot)
tissuespec_protein_brain = fix.read.csv(tissuespec_protein_brain)
tissuespec_protein_liver = fix.read.csv(tissuespec_protein_liver)
ts_protein_brain = fix.read.csv(ts_protein_brain)
ts_protein_liver = fix.read.csv(ts_protein_liver)
tsns_protein_brain = fix.read.csv(tsns_protein_brain)
tsns_protein_liver = fix.read.csv(tsns_protein_liver)

@@

A Wilcoxon-Mann-Whitney test shows no difference between the two peaks on either graph with p-values `r coin::pvalue(wil.brain) ` for brain and `r coin::pvalue(wil.liver)` for liver. 
We can try comparing the Tissue-Specific Secreted to the Ssec values of all of the origin tissue proteins. I need to create that dataset in the server; attempts to complete in chunks have been unsuccessful.

@@

```{r, eval = FALSE, echo = FALSE, warning = FALSE}

@@

>>>>

```{r echo=FALSE, message=FALSE}

@@

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_liver = read.csv("Liver_WGCNA.csv")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")
notsecreted_protein_brain = read.csv("NotSecreted_Brain.csv")
notsecreted_protein_liver = read.csv("NotSecreted_Liver.csv")
sec_annot = read.csv("Sec_Annotation.csv")
tissuespec_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
tissuespec_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
ts_protein_brain = read.csv("TissueSpecific2_Protein_Brain.csv")
ts_protein_liver = read.csv("TissueSpecific2_Protein_Liver.csv")
tsns_protein_brain = read.csv("NSTissueSpecific_Protein_Brain.csv")
tsns_protein_liver = read.csv("NSTissueSpecific_Protein_Liver.csv")

@@

protein_liver = fix.read.csv(protein_liver)
protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)
notsecreted_protein_brain = fix.read.csv(notsecreted_protein_brain)
notsecreted_protein_liver = fix.read.csv(notsecreted_protein_liver)
sec_annot = fix.read.csv(sec_annot)
tissuespec_protein_brain = fix.read.csv(tissuespec_protein_brain)
tissuespec_protein_liver = fix.read.csv(tissuespec_protein_liver)
ts_protein_brain = fix.read.csv(ts_protein_brain)
ts_protein_liver = fix.read.csv(ts_protein_liver)
tsns_protein_brain = fix.read.csv(tsns_protein_brain)
tsns_protein_liver = fix.read.csv(tsns_protein_liver)

@@

```{r}
# checking secreted

checking = sec_annot[sec_annot$ensembl_gene_id == top_brain[4],] # extracellular matrix

# ENSRNOG00000005770: "The WNT signalling modulator, Wise, is expressed in an interaction-dependent manner during hair-follicle cycling."

```
A Wilcoxon-Mann-Whitney test shows no difference between the two peaks on either graph with p-values `r coin::pvalue(wil.brain) ` for brain and `r coin::pvalue(wil.liver)` for liver. 
We can try comparing the Tissue-Specific Secreted to the Ssec values of all of the origin tissue proteins. I need to create that dataset in the server; attempts to complete in chunks have been unsuccessful.

@@

```{r}
# annotation

anno_brain = sec_annot[sec_annot$ensembl_gene_id %in% top_brain,]
anno_liver = sec_annot[sec_annot$ensembl_gene_id %in% top_liver,]

```
```{r, eval = FALSE, echo = FALSE, warning = FALSE}

@@

```{r, echo = FALSE}

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

full_brain = read.csv("Full_Brain_asso_TS.TSNS_Ssec.csv")
full_liver = read.csv("Full_Liver_asso_TS.TSNS_Ssec.csv")

# Visualizations
#hist(Ssec_brain$Ssec)
#hist(ts_Ssec_brain$Ssec)

# Ssec distribtions TSNS v. TS, brain and liver
#Brain
h1 = cbind(rep("Full brain", times = length(full_brain$Ssec)), full_brain$Ssec)
h2 = cbind(rep("TSS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)

ggplot(compare.b, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Brain Ssec Distribution: Full v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

# Liver
h1 = cbind(rep("Full liver", times = length(full_liver$Ssec)), full_liver$Ssec)
h2 = cbind(rep("TSS liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)

ggplot(compare.l, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Liver Ssec Distribution: Full v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

```
# compare secreted to non-secreted

```{r}

#### Secreted ####
## BRAIN ##
target_liver = cbind(ts_protein_liver, tsns_protein_liver)

brain.liv = bicorAndPvalue(secreted_protein_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

sec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
top_brain = head(sec_Ssec_brain[,1])
bot_brain = tail(sec_Ssec_brain[,1])

liv.brain = bicorAndPvalue(secreted_protein_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

sec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
top_liver = head(sec_Ssec_liver[,1])
bot_liver = tail(sec_Ssec_liver[,1])

#### Not Secreted ####

## BRAIN ##

brain.liv.c = bicorAndPvalue(notsecreted_protein_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

nsec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

liv.brain.c = bicorAndPvalue(notsecreted_protein_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

nsec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

####### Visualizations

#Brain
h1 = cbind(rep("Not Secreted brain", times = length(nsec_Ssec_brain$Ssec)), nsec_Ssec_brain$Ssec)
h2 = cbind(rep("Secreted brain", times = length(sec_Ssec_brain$Ssec)), sec_Ssec_brain$Ssec)

# Liver
h1 = cbind(rep("Not Secreted liver", times = length(nsec_Ssec_liver$Ssec)), nsec_Ssec_liver$Ssec)
h2 = cbind(rep("Secreted liver", times = length(sec_Ssec_liver$Ssec)), sec_Ssec_liver$Ssec)

wil.brain = coin::wilcox_test(val ~ trt, compare.b)
wil.brain

wil.liver = coin::wilcox_test(val ~ trt, compare.l)
wil.liver

# :'(

```{r}

# random brain secreted genes

sample(secreted_protein_brain, size = 4, replace = FALSE)

#ENSRNOG00000012655: extracellular matrix
# ENSRNOG00000053272: endo reticulum, extracellular space, cytoplasm, perinuclear region

checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000012655",]
checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000001806",]
checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000008519",]

checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000021020",]

# random liber secreted genes

sample(secreted_protein_liver, size = 4, replace = FALSE)

# ENSRNOG00000058444: nucleus, chromosome, secreted
# ENSRNOG00000048642: same

### 2021-10-07 Tissue Specific Report.Rmd
havent committed in a while

The ensembl rat liver and brain RNAseq data was acquired (get more info from papers) downloaded from PhenoGen. Inbred strains were excluded, leaving only recombinant inbred strains. Annontation data was retrieved from BiomaRt (version number 2.48.0), which was used to exclude non-protein-coding genes from the data set. This resulted in a dataset of 30 strains and 16,457 genes for brain tissue, and 30 strains and 14,109 genes for liver tissue.

From there, our datasets were segregated by terms secreted and tissue-specific. Results for protein-coding genes were filtered by species "Rattus norvegicus" and subcellular location "secreted" from the uniprot database (accessed 2021). The uniprot IDs from the new secreted dataset were compared to the converted uniprot IDs (via biomaRt) of the brain and liver tissue datasets. Tissue-specific status was determined by finding the genes in the brain and liver datasets were exclusive to those datasets (only found in the brain dataset OR the liver dataset). From this we created our dataset of investigation: tissue-specific secreted gene datasets for liver (with 35 genes) and brain (with 174 genes) tissues. This dataset was compared to a comparable dataset of tissue-specific non-secreted genes, with _ liver genes and _ brain genes.

@@

# Data Input
setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")

@@

@@

## Tissue Specific Non-Secreted
setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
notsec_brain = read.csv("NotSecreted_Brain.csv")
notsec_liver = read.csv("NotSecreted_Liver.csv")
notsec_liver = fix.read.csv(notsec_liver)
notsec_brain = fix.read.csv(notsec_brain)

@@

knitr::kable(num, caption = "Number of Genes", col.names = NA, )

# compare shared genes between brain and liver

gee = intersect(colnames(ts_protein_brain), colnames(ts_protein_liver))

int = c(length(colnames(ts_protein_brain)), length(gee), length(colnames(ts_protein_liver)))
names(int) = c("Tissue-Specific Secreted Genes Brain", "Tissue-Specific Secreted Genes Shared by Brain and Liver", "Tissue-Specific Secreted Genes Liver")

knitr::kable(int)

hee = intersect(colnames(tsns_protein_brain), colnames(tsns_protein_liver))

int2 = c(length(colnames(tsns_protein_brain)), length(hee), length(colnames(tsns_protein_liver)))
names(int2) = c("Tissue-Specific Non-Secreted Genes Brain", "Tissue-Specific Non-Secreted Genes Shared by Brain and Liver", "Tissue-Specific Non-Secreted Genes Liver")

knitr::kable(int2)

@@

wil.b = rbind(gee, gee2)
wil.b = as.data.frame(wil.b)
colnames(wil.b) = c("trt", "r.sq")

wil.b$trt = as.factor(wil.b$trt)
wil.b$r.sq = as.numeric(wil.b$r.sq)

wil.brain = coin::wilcox_test(r.sq ~ trt, wil.b)

@@

wil.l = rbind(gee, gee2)
wil.l = as.data.frame(wil.l)
colnames(wil.l) = c("trt", "r.sq")

wil.l$trt = as.factor(wil.l$trt)
wil.l$r.sq = as.numeric(wil.l$r.sq)

wil.liver = coin::wilcox_test(r.sq ~ trt, wil.l)

@@

ggplot(wil.b, aes(x = r.sq, color = trt)) + 
  geom_density() +
  labs(title="R Squared Density Curve: Brain",x="R Squared", y = "Density")

ggplot(wil.l, aes(x = r.sq, color = trt)) + 
  geom_density() +
  labs(title="R Squared Density Curve: Liver",x="R Squared", y = "Density")

@@

boxplot(r.squared_brain_tsns[,2], r.squared_brain_ts[,2], names = c("Tissue-Specific Non-Secreted", "Tissue-Specific Secreted"), xlab = "R Squared", main = "R Squared Brain Data")
boxplot(r.squared_liver_tsns[,2], r.squared_liver_ts[,2], names = c("Tissue-Specific Non-Secreted", "Tissue-Specific Secreted"), xlab = "R Squared", main = "R Squared Liver Data")

@@

We tested whether the heritability of the Tissue-Specific Non-Secreted and Tissue-Specific Secreted datasets were significantly different using the Wilcoxon-Mann-Whitney Test. Only the liver dataset showed a significant difference betweeen the R Squared of the two groups.

Tissue-Specific Non-Secreted Brain Dataset v. Tissue-Specific Secreted Brain Dataset:

```{r, echo = FALSE, message = FALSE}
wil.brain

Tissue-Specific Non-Secreted Liver Dataset v. Tissue-Specific Secreted Liver Dataset:

wil.liver

>>>>

The ensembl rat liver and brain RNAseq data was produced by our lab from tissues imported from the Czech Academy of Sciences laboratory. Inbred strains were excluded, leaving only recombinant inbred strains. The means across strains were calculated for the RNA expression data. Annontation data was retrieved from BiomaRt (version number 2.48.0), which was used to exclude non-protein-coding genes from the data set. This resulted in a dataset of 30 strains and 16,457 genes for brain tissue, and 30 strains and 14,109 genes for liver tissue.

From there, our datasets were segregated by terms "secreted" and "tissue-specific." Results for protein-coding genes were filtered by species "Rattus norvegicus" and subcellular location "secreted" from the uniprot database (accessed 2021). The uniprot IDs from the new secreted dataset were compared to the converted uniprot IDs (via biomaRt) of the brain and liver tissue datasets. Tissue-specific status was determined by finding the genes in the brain and liver datasets were exclusive to those datasets (only found in the brain dataset OR the liver dataset). From this we created our dataset of investigation: tissue-specific secreted gene datasets for liver (with 35 genes) and brain (with 174 genes) tissues. This dataset was compared to a comparable dataset of tissue-specific non-secreted genes, with 368 liver genes and 2,577 brain genes.

@@

# Data Input
#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")

@@

# flow chart
# https://towardsdatascience.com/r-visualizations-flow-charts-in-r-4cfa7f783872
library(DiagrammeR)

grViz(diagram = "digraph flowchart { 
      node [fontname = Arial, shape = oval, color = Lavender, style = filled]
      tab1 [label = '@@1', fontsize = 100]
      tab2 [label = '@@2', fontsize = 100]
      tab3 [label = '@@3', fontsize = 100]
      tab4 [label = '@@4', fontsize = 100]
      tab5 [label = '@@5', fontsize = 100]
      tab6 [label = '@@6', fontsize = 100]
      tab7 [label = '@@7', fontsize = 100]
      tab8 [label = '@@8', fontsize = 100]

      edge[minlen = 4, penwidth = 8, arrowsize = 4]
      tab1 -> tab2;
      tab2 -> tab3;
      tab3 -> tab4;
      tab4 -> tab5;
      tab5 -> tab6;
      tab4 -> tab7;
      tab7 -> tab8}

      [1]: 'Full Cleaned Data: Brain (132 individuals, 18,385 genes), Liver (132 individuals, 15,365 genes)'
      [2]: 'Means of Strains: Brain (45 strains, 18,385 genes), Liver (45 strains, 15,365 genes)'
      [3]: 'RI Strains Only: Brain (30 strains, 18,385 genes), Liver (30 strains, 15,365 genes)'
      [4]: 'Protein-Coding Genes Only: Brain (30 strains, 16,457 genes), Liver (30 strains, 14,109 genes)'
      [5]: 'Secreted Protein-Coding Genes: B(30, 657), L(30, 818)'
      [6]: 'Tissue-Specific Secreted Protein-Coding Genes: B(30, 174), L(30, 35)'
      [7]: 'Non-Secreted Protein-Coding Genes: B(30, 15,800), L(30, 13,591)'
      [8]: 'Tissue-Specific Non-Secreted Protein-Coding Genes: B(30, 2,577), L(30, 368)' 
      ", width = 1000, height = 400)

@@

## Tissue Specific Non-Secreted
#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
notsec_brain = read.csv("NotSecreted_Brain.csv")
notsec_liver = read.csv("NotSecreted_Liver.csv")
notsec_liver = fix.read.csv(notsec_liver)
notsec_brain = fix.read.csv(notsec_brain)

@@

knitr::kable(num, col.names = "Number of Ensembl Genes")

@@

wil.b = rbind(gee, gee2)
wil.b = as.data.frame(wil.b)
colnames(wil.b) = c("Dataset", "r.sq")

wil.b$Dataset = as.factor(wil.b$Dataset)
wil.b$r.sq = as.numeric(wil.b$r.sq)

wil.brain = coin::wilcox_test(r.sq ~ Dataset, wil.b)

@@

wil.l = rbind(gee, gee2)
wil.l = as.data.frame(wil.l)
colnames(wil.l) = c("Dataset", "r.sq")

wil.l$Dataset = as.factor(wil.l$Dataset)
wil.l$r.sq = as.numeric(wil.l$r.sq)

wil.liver = coin::wilcox_test(r.sq ~ Dataset, wil.l)

@@

ggplot(wil.b, aes(x = r.sq, color = Dataset)) + 
  geom_density() +
  labs(title ="Broad Sense Heritability Density Curve: Brain",x = "R Squared", y = "Density")

ggplot(wil.l, aes(x = r.sq, color = Dataset)) + 
  geom_density() +
  labs(title ="Broad Sense Heritability Density Curve: Liver",x = "R Squared", y = "Density")

@@

boxplot(r.squared_brain_tsns[,2], r.squared_brain_ts[,2], names = c("Brain-Specific Non-Secreted", "Brain-Specific Secreted"), main = "Broad Sense Heritability: Brain", ylab = "R Squared")
boxplot(r.squared_liver_tsns[,2], r.squared_liver_ts[,2], names = c("Liver-Specific Non-Secreted", "Liver-Specific Secreted"), main = "Broad Sense Heritability: Liver", ylab = "R Squared")

@@

We tested whether the heritability of the Tissue-Specific Non-Secreted and Tissue-Specific Secreted datasets were significantly different using the Wilcoxon-Mann-Whitney Test. Only the liver dataset showed a significant difference betweeen the R Squared of the two groups. The test comparing the Brain-Specific Non-Secreted and Brain-Specific Secreted resulted in a p-value of `r round(coin::pvalue(wil.brain), 2)`. The test comparing the Liver-Specific Non-Secreted and Liver-Specific Secreted resulted in a p-value of `r format(coin::pvalue(wil.liver), scientific = TRUE, digits = 3)`.
